---
title: 'Core Module 3 Week 2: Part 2: Research Question'
author: "Jane Ngala"
date: '2022-06-04'
output: html_document
---

**1. Defining the Question**

***(a) Specifying the Question***

- To learn and understand the characteristics of customer groups

***(b) Metric of Success***

- The project will be considered a success if findings on the characteristics of customer groups helps inform the team in formulating the marketing and   sales strategies of the brand

***(c) Understanding the context***

- Kira Plastinina is a Russian brand that is sold through a defunct chain of retail stores in Russia, Ukraine, Kazakhstan, Belarus, China, Philippines, and Armenia. The brand’s Sales and Marketing team would like to understand their customer’s behavior from data that they have collected over the past year. More specifically, they would like to learn the characteristics of customer groups.

***(d) Experimental Design***

- Loading libraries
- Loading Dataset
- Data Cleaning
- EDA
- Implement the solution
- Challenge the Solution
- Follow up Questions

***(e) Appropriateness of Data***

- Data provided is appropriate and reliable

**2. Loading Libraries and Dataset**

```{r}
# Loading libraries
library(data.table)
library (plyr)
library(ggplot2)
library(moments)
library(ggcorrplot)

```

```{r}
# Loading Dataset

url<-"http://bit.ly/EcommerceCustomersDataset"
kira<-read.csv(url)
```


**3. Checking the Data**

```{r}
# Previewing the data

#view(kira)
```

```{r}
# Previewing top of the data

head(kira)
```

```{r}
# Previewing bottom of the data

tail(kira)
```

```{r}
# Previewing shape

cat("The dataset has", nrow(kira), "rows", "and", ncol(kira), "columns")
```

```{r}
# Checking Data types

str(kira)
```
- All columns have appropriate data types

**4. Data Cleaning**

```{r}
# Checking for number of missing values

length(which(is.na(kira)))
```
- There are 112 missing values.
- Taking a closer look at the missing data.

```{r}
# Displayng the missing values

kira[!complete.cases(kira), ]
```
- Missing values are in the 1st eight columns, all are numerical columns.

```{r}
# Handling missing values using the mean

kira$Administrative[is.na(kira$Administrative)] <- mean(kira$Administrative, na.rm = TRUE)
kira$Administrative_Duration[is.na(kira$Administrative_Duration)] <- mean(kira$Administrative_Duration, na.rm = TRUE)
kira$Informational[is.na(kira$Informational)] <- mean(kira$Informational, na.rm = TRUE)
kira$Informational_Duration[is.na(kira$Informational_Duration)] <- mean(kira$Informational_Duration, na.rm = TRUE)
kira$ProductRelated[is.na(kira$ProductRelated)] <- mean(kira$ProductRelated, na.rm = TRUE)
kira$ProductRelated_Duration[is.na(kira$ProductRelated_Duration)] <- mean(kira$ProductRelated_Duration, na.rm = TRUE)
kira$BounceRates[is.na(kira$BounceRates)] <- mean(kira$BounceRates, na.rm = TRUE)
kira$ExitRates[is.na(kira$ExitRates)] <- mean(kira$ExitRates, na.rm = TRUE)
sum(is.na(kira))
```

```{r}
# Checking for duplicates

sum(duplicated(kira))
```
- There are 119 duplicated records in the dataset

```{r}
# Removing duplicates

kira <- kira[!duplicated(kira), ]
sum(duplicated(kira))
```


```{r}
# Tidying column names

colnames(kira)
```
- Column names are well structured and have similar case.

```{r}
# Checking for outliers
# Selecting numerical columns

cols_num  <- unlist(lapply(kira, is.numeric))
kira_num <- kira[ ,cols_num]
boxplot(kira_num)
```

- The columns with outliers are: Administrative_Duration,Informational_Duration and ProductRelated_Duration

```{r}
# Finding total number of outliers

sum(kira$Administrative_Duration < quantile(kira$Administrative_Duration, p = 0.25)- 1.5 * IQR(kira$Administrative_Duration)) + sum(kira$Administrative_Duration > quantile(kira$Administrative_Duration, p = 0.75) + 1.5 * IQR(kira$Administrative_Duration))

sum(kira$Informational_Duration < quantile(kira$Informational_Duration, p = 0.25)- 1.5 * IQR(kira$Informational_Duration)) + sum(kira$Informational_Duration > quantile(kira$Informational_Duration, p = 0.75) + 1.5 * IQR(kira$Informational_Duration))

sum(kira$ProductRelated_Duration < quantile(kira$ProductRelated_Duration, p = 0.25)- 1.5 * IQR(kira$ProductRelated_Duration)) + sum(kira$ProductRelated_Duration > quantile(kira$ProductRelated_Duration, p = 0.75) + 1.5 * IQR(kira$ProductRelated_Duration))
```
- Outliers are too many, if removed there will be insufficient data for analysis

```{r}
# Previewing dimensions of the clean data

dim(kira)
```


**5. EDA**

```{r}
# Summary of statistical distribution

summary(kira)
```

  ***(i) Univariate Analysis***

```{r}
# Visualization using barplots

par(mfrow = c(2,4), mar = c(4,3,2,2))
barplot(table(kira$Revenue),main = "Revenue")
barplot(table(kira$Weekend),main = "Weekend")
barplot(table(kira$VisitorType),main = "Visitor Type")
barplot(table(kira$TrafficType),main = "Traffic Type")
barplot(table(kira$Region),main = "Region")
barplot(table(kira$Browser),main = "Browser")
barplot(table(kira$OperatingSystems),main = "Operating Systems")
barplot(table(kira$Month),main = "Month")


```

- Most customers visit the site during weekdays
- Most customers visiting the site do not purchase
- Most visitors on the site are returning customers
- The most popular operating system is OS 2
- The most popular region is region 1 followed by region 3
- The most popular browser used by customers is browser 2 followed by browser 1

```{r}
# Visualization using Histograms

par(mfrow = c(2,2), mar = c(5,4,2,2))
hist(kira$Administrative,xlab ='Administrative', main ='Histogram for Administrative')
hist(kira$Administrative_Duration,xlab ='Administrative Duration', main ='Histogram for Administrative Duration')
hist(kira$Informational,xlab ='Informational', main ='Histogram for Informational')
hist(kira$Informational_Duration,xlab ='Informational Duration', main ='Histogram for Informational Duration')
hist(kira$ProductRelated,xlab ='Product Related', main ='Histogram for Product Related')
hist(kira$ProductRelated_Duration,xlab ='Product Related Duration', main ='Histogram for Product Related Duration')
hist(kira$BounceRates,xlab ='Bounce Rates', main ='Histogram for Bounce Rates')
hist(kira$ExitRates,xlab ='Exit Rates', main ='Histogram for Exit Rates')
hist(kira$PageValues,xlab ='Page Values', main ='Histogram for Page Values')
hist(kira$SpecialDay,xlab ='Special Day', main ='Histogram for Special Day')
```

- Data is heavily skewed to the left for the numerical columns;Distribution is not Gaussian.

  ***(ii) Bivariate Analysis***

```{r}
# Visualization of relationship between Admin duration and product related duration

ggplot(kira, aes(x=ProductRelated_Duration, y=Administrative_Duration)) + geom_point()

```

```{r}
# Visualization of relationship between Administrative and Administrative duration

ggplot(kira, aes(x=Administrative, y=Administrative_Duration)) + geom_point()
```

```{r}
# Visualization of relationship between Informational and Informational duration

ggplot(kira, aes(x=Informational, y=Informational_Duration)) + geom_point()
```

```{r}
# Visualization of relationship between Product Related and Product related duratiion

ggplot(kira, aes(x=ProductRelated, y=ProductRelated_Duration)) + geom_point()
```


```{r}
# Plotting the correlations

ggcorrplot(cor(kira[ , c(1:10)]), type = "lower", outline.col = "black",
 lab=TRUE,
 ggtheme = ggplot2::theme_gray,
 colors = c("#6D9EC1", "white", "#E46726"))
```
- A positive correlation indicates the variables have a linear relationship
- The variables in our data that have a high positive correlation are:
      * ExitRates & BounceRates
      * ProductRelated & ProductRelated_Duration 
      * Informational &  Informational_Duration
      * Administrative & Administartive_Duration 


**6. Implementing the Solution**

```{r}
# Dropping the Character variables 

kira_new<-kira[ ,c(1:10,12:15)]
```


```{r}
# Normalizing the data

normalize <- function(x){return ((x-min(x)) / (max(x)-min(x)))}

kira_norm<-normalize(kira_new)

```

***(i) K-Means Clustering***

```{r}
# Loading Libraries

library(factoextra)
library(NbClust)
library (cluster)


```


```{r}
# Elbow method

fviz_nbclust(kira_norm, kmeans, method = "wss") + geom_vline(xintercept = 3, linetype = 2)+ labs(subtitle = "Elbow method")
```
- Optimum number of clusters is 3 using elbow method

```{r}
# Silhouette method

fviz_nbclust(kira_norm, kmeans, method = "silhouette")+ labs(subtitle = "Silhouette method")
```
- Optimum number of clusters is 2 using silhouette method

```{r}
# Choosing the best number of clusters

# nb<-NbClust(data = kira_norm, distance = "euclidean",min.nc = 2, max.nc = 15, method = "kmeans")
# fviz_nbclust(nb)
```

```{r}
# Implementing  using 2 clusters

km <- kmeans(kira_norm,2,iter.max = 10, nstart = 25)
km
```

```{r}
# Plotting  final k-means model
 
fviz_cluster(km, data = kira_norm)
```

***(ii) Hierarchical Clustering***

```{r}
# Using randomization

rand_kira <- kira_norm[sample(nrow(kira_norm), size=1500), ]
head(rand_kira)


```


```{r}
# Defining linkage methods

methods<- c( "average", "single", "complete", "ward")
names(methods) <- c( "average", "single", "complete", "ward")

```

```{r}
# Function to compute agglomerative coefficient

ac <- function(x) {agnes(rand_kira, method = x)$ac}
```

```{r}
# Calculating agglomerative coefficient for each clustering linkage method

sapply(methods, ac)
```

```{r}
# Performing hierarchical clustering using Ward's minimum variance

clust <- agnes(rand_kira, method = "ward")
```

```{r}
# Producing dendrogram

pltree(clust, cex = 0.6, hang = -1, main = "Dendrogram") 
```

```{r}
# Calculating gap statistic 

gap_stat <- clusGap(rand_kira, FUN = hcut, nstart = 25, K.max = 10, B = 50)
```

```{r}
# Plotting clusters vs. gap statistic
fviz_gap_stat(gap_stat)

```

```{r}
# Applying Cluster Labels to dataset
# Computing distance matrix
# Performing hierarchical clustering using Ward's method

final_clust <- agnes(rand_kira,metric = "euclidean", method = "ward",stand = TRUE)

# cutting the dendrogram into 4 clusters

groups <- cutree(final_clust, k=3)

# Finding number of observations in each cluster
table(groups)

```

```{r}
# Appending cluster labels to original data

final_data <- cbind(rand_kira, cluster = groups)
 
# Displaying the top of the final data

head(final_data)
```

```{r}
# Finding mean values for each cluster

aggregate(final_data, by=list(cluster=final_data$cluster), mean)
```

**7. Conclusion**
 
- Most customers visit the site during weekdays
- Most customers visiting the site do not make a purchase
- Most visitors on the site are returning customers
- The most popular operating system is OS 2
- The most popular region is region 1 followed by region 3
- The most popular browser used by customers is browser 2 followed by browser 1
  
